---
title: "Figures for SAPFLUXNET data paper"
author: "R. Poyatos"
date: "27/12/2019"
output:
  word_document: 
    reference_docx: word-styles-reference-01.docx
  html_document:
    df_print: paged
  pdf_document: 
    fig_caption: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sapfluxnetr)
library(purrr)
library(flextable)
library(pander)
library(magrittr)
library(ggplot2)
library(ggalluvial)
library(ggrepel)
library(ggbiome)
library(cowplot)
library(ggspatial)
library(raster)

source('R/sfn_datapaper_functions.R')
source('metadata_wrangling.R')
load('maps_base.RData')

```

## Figures

### Figure 1. 
- Data workflow. Víctor is on it...

### Figure 2
- This version of the figures uses a low quality version of the base maps because 
it's faster to run this way. Inaccuracies in the map result from rasterisation.
- I include two versions of the figure with the maps, one with numeric codes and 
another without them. Need to improve the layout and/or tweak the parameters controlling 
the position of the labels to get an optimal figure.
- TODO: use paler colour for forest area, see if I can display IGBP code (EBF, ENF, etc)

### Figure 3. Biomes.
- This figure is OK.
- TODO: explore the possibility that Fig. 2 is (a) world map, (b) biome plot and that figure (3)
is more regionally detailed, showing PFT.

### Figure 4. Taxonomy.
- I chose to show genera with >50 trees and species with >30 trees, arbitrarily...
- a) add number of species at the tip
- b) add number of sites at the tip
- Add a third panel ranking species by tree·days?

### Figure 5. Sap flow methods and level.
- OK
### Figure 6. Site and plant level attributes
- TODO: add titles to labels a) stand, b) plant

### Figure 7. Datasets duration and period
- TODO: improve with Víctor.

### Figure 8. Temporal patterns
- TODO: add titles to labels: a) ESP_CAN, Mediterranean, b), c

\pagebreak

```{r data_workflow, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
knitr::include_graphics('resources/QC_process.svg')
```

Figure 1. Data workflow.

\pagebreak

```{r sitesmap, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, fig.height=15, fig.width=15}
sitesmap_global <- globforest_rec_0_1  +
  geom_point(data=sfn_allsites,aes(x=si_long,y=si_lat),shape=21,color='black',fill='royalblue',alpha=0.5)+
  guides(fill='none')+xlab(NULL)+ylab(NULL)+
  theme_light()+
  theme(axis.title=element_text(size=18),axis.text=element_text(size=18))

# TODO: preliminary filtering out climatic NAs
# Use CHELSA to gf 

biomes_plot <- sfn_allsites %>%
  dplyr::arrange(si_biome) %>%
  dplyr::filter(!is.na(si_map) | !is.na(si_map)) %>% 
  ggbiome::vis_location_biome(si_lat='si_lat',
                              si_long='si_long',si_mat='si_mat',si_map='si_map') +
  
  theme(
    panel.grid = element_blank(),
    axis.ticks = element_blank(),
    title= element_text(size=20),
    axis.title=element_text(size=20),
    axis.text=element_text(size=20),
    legend.title = element_blank(),
    legend.text = element_text(size =16),
    legend.key.size = unit(1.5, 'line'),
    legend.position = c(0.7,0.2)
    
  )

igbp_plot<-sfn_allsites %>% 
  group_by(si_igbp) %>% 
  tally() %>% 
  ggplot(.,aes(y=n,x=reorder(si_igbp,desc(n))))+
  geom_bar(stat='identity')+
  xlab('Vegetation type')+ylab('Number of datasets')+
  theme_light()+
  theme(axis.title=element_text(size=18),axis.text=element_text(size=18))

# Build figure 
cowplot::plot_grid(
  # row 1
  cowplot::plot_grid(sitesmap_global, labels='(a)'),
  # row 2
  cowplot::plot_grid(biomes_plot, igbp_plot,labels=c('(b)','(c)'),rel_widths=c(0.6,0.4)),
  ncol=1, label_size=16, rel_heights=c(0.6,0.4))

```

Figure 2. Geographical distribution of SAPFLUXNET datasets.

\pagebreak


```{r regional_maps,fig.height=18, fig.width=15, cache=TRUE,echo=FALSE, message=FALSE, warning=FALSE,}

# a) Site labels 

# b) Select countries to plot 

countries_sfn <- unique(sapply(strsplit(sfn_allsites$si_code,"_"),"[[",1))
countries_label <- c('ARG','BRA','COL','CRI','CHN','GUF','IDN','ISR','KOR','MDG','MEX','NZL','RUS','SEN','THA','UZB')
countries_europe <- c('AUT','CHE','CZE','DEU','ESP','FIN','FRA','GBR','HUN','ITA','NLD','PRT','SWE')
countries_america <- c('CAN','USA')
countries_austral <- c('AUS')
countries_africa <- c('MDG','ZAF')


# c) Add numeric codes 

sfn_allsites_country<- sfn_allsites %>% 
  mutate(country=sapply(strsplit(si_code,"_"),"[[",1),
         num_code = as.integer(factor(si_code)))


# 3. Map, numeric codes, ggrepel not aligned ------------------------------------------------------

# world
sfnsites_world_dr <- globforest_rec_0_1  +
  geom_point(data=sfn_allsites,aes(x=si_long,y=si_lat),shape=21,color='black',fill='royalblue',alpha=0.5)+
  geom_text_repel(data = subset(sfn_allsites_country,country%in%countries_label), 
                  aes(si_long, si_lat, label = num_code), size = 3,
                  segment.alpha=0.5,  nudge_x = -0.35,direction = "both",
                  box.padding = unit(0.1, 'lines'), force = 0.5)+geom_label_repel()+guides(fill='none')+xlab(NULL)+ylab(NULL)

# europe
sfnsites_europe_dr<- globforest_rec_0_1  +
  geom_point(data=sfn_allsites,aes(x=si_long,y=si_lat),shape=21,color='black',fill='royalblue',alpha=0.5)+
  geom_text_repel(data = subset(sfn_allsites_country,country%in%countries_europe), 
                  aes(si_long, si_lat, label = num_code), size = 3,
                  segment.alpha=0.5,  nudge_y = -0.35,direction = "both",
                  box.padding = unit(0.1, 'lines'), force = 0.5)+
  coord_sf(xlim = c(-20, 40), ylim = c(32, 67), expand = FALSE)+
  geom_label_repel()+guides(fill='none')+xlab(NULL)+ylab(NULL)

# america
sfnsites_america_dr<- globforest_rec_0_1  +
  geom_point(data=sfn_allsites,aes(x=si_long,y=si_lat),shape=21,color='black',fill='royalblue',alpha=0.5)+
  geom_text_repel(data = subset(sfn_allsites_country,country%in%countries_america), 
                  aes(si_long, si_lat, label = num_code), size = 3,
                  segment.alpha=0.5,  nudge_x = -0.35,direction = "both",
                  box.padding = unit(0.1, 'lines'), force = 0.5)+
  coord_sf(xlim = c(-130, -60), ylim = c(30, 55), expand = FALSE)+
  geom_label_repel()+guides(fill='none')+xlab(NULL)+ylab(NULL)

# austral
sfnsites_austral_dr<- globforest_rec_0_1  +
  geom_point(data=sfn_allsites,aes(x=si_long,y=si_lat),shape=21,color='black',fill='royalblue',alpha=0.5)+
  geom_text_repel(data = subset(sfn_allsites_country,country%in%countries_austral), 
                  aes(si_long, si_lat, label = num_code), size = 3,
                  segment.alpha=0.5,  nudge_x = -0.35,
                  box.padding = unit(0.1, 'lines'), force = 0.5)+
  coord_sf(xlim = c(110, 162), ylim = c(-10, -50), expand = FALSE)+
  geom_label_repel()+guides(fill='none')+xlab(NULL)+ylab(NULL)

# africa
sfnsites_africa_dr<- globforest_rec_0_1  +
  geom_point(data=sfn_allsites,aes(x=si_long,y=si_lat),shape=21,color='black',fill='royalblue',alpha=0.5)+
  geom_text_repel(data = subset(sfn_allsites_country,country%in%countries_africa), aes(si_long, si_lat, label = num_code), size = 3,
                  box.padding = unit(0.1, 'lines'), force = 0.5)+
  coord_sf(xlim = c(10, 50), ylim = c(-10, -40), expand = FALSE)+
  geom_label_repel()+guides(fill='none')+xlab(NULL)+ylab(NULL)



# e) Build figure 

plot_grid(
  plot_grid(sfnsites_world_dr,ncol=1,nrow=1,labels=c('a)')),
  plot_grid(sfnsites_europe_dr,sfnsites_austral_dr,ncol=2,nrow=1,rel_widths=c(1,0.7),labels=c('b)','c)')),
  plot_grid(sfnsites_america_dr,sfnsites_africa_dr,ncol=2,nrow=1,rel_widths=c(1,0.7),labels=c('d)','e)')),
  nrow=3,rel_heights=c(1,1,0.7))

```
Figure 3. Detailed geographic distribution.

\pagebreak

```{r genera_species,fig.height=10, fig.width=15, echo=FALSE}

sfn_sitespecies_tax %>%
  group_by(genus,si_code) %>%
  summarise(n_si = sum(sp_ntrees)) %>% 
  group_by(genus) %>% 
  mutate(
    n_all = sum(n_si),
    vjust = n_si/sum(n_si),
    fill_dummy = if_else(n_si > 14, si_code, 'zz_other')
  ) %>%
  filter(n_all>50) %>%
  arrange(desc(n_si)) %>%
  mutate(
    accumulated = cumsum(n_si),
    segment_start = accumulated - (n_si/2)
  ) -> ntrees_species_top_data

ntrees_species_top_data %>%
  # only show label for those sites in each genus with 15 trees or more
  filter(n_si > 14) -> ntrees_species_top_data_labels

ntrees_species_top <- ntrees_species_top_data %>% 
  ggplot(aes(x = reorder(genus, n_all), y = n_si, fill = reorder(fill_dummy, n_si))) +
  geom_bar(stat = 'identity', show.legend = FALSE) +
  geom_text(
    aes(label = si_code, y = segment_start),
    data = ntrees_species_top_data_labels,
    # direction = 'both', min.segment.length = 0.3,
    # arrow = arrow(length = unit(0.01, 'npc'), type = 'open'),
    # hjust = 0.75,
    # vjust = 1.5,
    angle = 270, size = 2, position = 'identity'
  ) +
  viridis::scale_fill_viridis(discrete = TRUE) +
  labs(x = '', y = 'Number of trees',
       title = 'Genera with > 50 trees',
       subtitle = '(label only showed for sites with >15 trees per genera)') +
  coord_flip() +
  theme_bw() +
  theme(
    title= element_text(size=16),
    axis.title=element_text(size=16),
    axis.text=element_text(size=16),
    axis.text.y=element_text(face='italic'),
    legend.title = element_blank(),
    legend.position = 'none'
  )


ntrees_genus_top <- sfn_sitespecies_tax%>%
  filter(!is.na(sp_ntrees)) %>% 
  group_by(sp_name, si_code) %>%
  summarise(n = sum(sp_ntrees)) %>%
  group_by(sp_name) %>% 
  mutate(n_all=sum(n)) %>% 
  filter(n_all>30) %>% 
  ggplot(aes(x = reorder(sp_name, n, sum), y = n, fill = si_code)) +
  geom_bar(stat = 'identity')+
  viridis::scale_fill_viridis(discrete = TRUE) +
  labs(x = '', y = 'Number of trees', 
       title = 'Species with >30 trees') +
  coord_flip() +
  theme_bw() +
  theme(
    title= element_text(size=16),
    axis.title=element_text(size=16),
    axis.text=element_text(size=16),
    axis.text.y=element_text(face='italic'),
    legend.title = element_blank(),
    legend.position = 'none'
  )

cowplot::plot_grid(ntrees_genus_top,ntrees_species_top,
                   label_size= 16,labels=c('a)', 'b)'),ncol=1, nrow=2, align = "v", axis = "b")
```
Figure 4. Taxonomic distribution of genera and species in SAPFLUXNET, showing (a) genera with > 50 trees and (b) species with > 30 trees in the database.

\pagebreak

```{r method_type,fig.height=18, fig.width=15, echo=FALSE}

ggplot(sfn_plants_type,
       aes(y = n, axis2 = pl_sens_meth, axis3= typef,axis1=group)) +
  geom_alluvium(aes(fill = group), width = 1/12)+
  geom_stratum(width = 1/12, alpha = .5) +
  geom_label_repel(stat = "stratum", infer.label = TRUE) +
  scale_x_discrete(limits = c("Group","Sap flow method", "Measurement level"), 
                   expand = c(.05, .05)) +
  #scale_fill_brewer(type = "qual", palette = "Set1")+
  scale_fill_viridis_d()+
  ylab('Number of trees')+
  theme_minimal()+
  theme(legend.position = "none", axis.title=element_text(size=16),
        axis.text = element_text(size=16))
  
```

Figure 5. Distribution of trees in SAPFLUXNET according to sap flow method and measurement level.

\pagebreak

```{r stand_plant_attributes, echo=FALSE, message=FALSE, warning=FALSE, fig.height=15, fig.width=18}

# Stand attributes --------------------------------------------------------

sfn_allstands %>%
  dplyr::select(si_code, st_age, st_height, st_density,
                        st_basal_area, st_lai, st_soil_depth) %>%
  pivot_longer(names_to="Var",values_to= "Value", cols=st_age:st_soil_depth)-> sfn_allstands_long

sfn_allstands_long$Varf <- factor(sfn_allstands_long$Var,
                                 levels=c('st_age','st_height','st_density','st_basal_area','st_lai','st_soil_depth'),
                                 labels=c('Age~(years)','Height~(m)','Stem~density~(stems~ha^{"-1"})',
                                          'Basal~area~(m^2~ha^{"-1"})','LAI~(m^2~m^{"-2"})','Soil~depth~(m)'))

sfn_allstands_long %>% 
  ggplot(.,aes(x = Var, y = Value)) +
  geom_violin(fill = '#FDE3A7', alpha = 0.3) +
  geom_point(colour='black',alpha=0.5,
             position = position_jitter(width = 0.2, height = 0))+
  facet_wrap(~Varf,scales='free',labeller=label_parsed)+
  guides(fill = FALSE) +
  labs(x=NULL,y=NULL)+
  theme_light()+
  theme(axis.text.x=element_blank(),axis.text.y=element_text(size=20),
        strip.background = element_rect(fill='white',colour='darkgray'),
        strip.text=element_text(size=20,colour='black',face='bold')) -> stand_violins

#  Plant attributes -------------------------------------------------------

sfn_allplants %>% 
  dplyr::select(pl_code,pl_dbh,pl_height,pl_sapw_area,pl_sapw_depth,pl_leaf_area) %>% 
  pivot_longer(names_to="Var", values_to="Value",cols=pl_dbh:pl_leaf_area)-> sfn_allplants_long

sfn_allplants_long$Varf <- factor(sfn_allplants_long$Var,
                                  levels=c('pl_dbh','pl_height','pl_sapw_area','pl_sapw_depth','pl_leaf_area'),
                                  labels=c('DBH~(cm)','Height~(m)','Sapwood~area~(cm^2)','Sapwood~depth~(cm)','Leaf~area(m^2)'))


sfn_allplants_long%>% 
  ggplot(.,aes(x = Var, y = Value)) +
  geom_violin(fill = '#FDE3A7', alpha = 0.3) +
  geom_point(colour='black',alpha=0.5,
             position = position_jitter(width = 0.2, height = 0))+
  facet_wrap(~Varf,scales='free',labeller=label_parsed)+
  guides(fill = FALSE) +
  labs(x=NULL,y=NULL)+
  theme_light()+
  theme(axis.text.x=element_blank(),axis.text.y=element_text(size=20),
        strip.background = element_rect(fill='white',colour='darkgray'),
        strip.text=element_text(size=20,colour='black',face='bold')) -> plant_violins

cowplot::plot_grid(stand_violins, plant_violins,labels=c('(a)','(b)'),rel_widths=c(1,1),nrow=2)

```


Figure 6. Plant and stand attributes

\pagebreak

```{r dataset_duration, echo=FALSE, cache=TRUE, message=FALSE, warning=FALSE, fig.height=18, fig.width=15 }
# Load data
load('dataset_duration_data.RData')

  # Figure showing dataset duration and n_trees
   
  # Palette for dataset duration plot
  treeclass_pal<- c('lightgrey',viridis::viridis_pal(direction=-1)(7))
  
  data_duration_plot<- 
    datasets_duration %>% 
    ggplot(., 
           aes( y=index_days,x=reorder(num_code,n),fill = n_trees_class_f)) +
    geom_tile()+
    coord_flip()+
    scale_fill_manual(values = treeclass_pal)+
    labs(x='',y='Number of days',fill='Number of trees')+
    theme_light()+
      theme(legend.position=c(.3,.2),
            axis.title.x=element_text(size=14),axis.text.x=element_text(size=12),
            axis.text.y=element_blank())+
            
      
    geom_label_repel(data=duration_annotation %>% 
                filter(rank%in%1:30),aes(x=reorder(num_code,n),y=0.9*n,label=si_code),inherit.aes=FALSE,
                ylim=c(3800,6000),direction='y',segment.color='darkgrey',force=0.65,size=3.4,label.size=0.03,
                
                # WARNING:If size is set to 3, the final plot only shows some, not all, labels ??
                
                nudge_x=duration_annotation %>% filter(rank%in%1:30) %>% arrange(desc(n)) %>% pull(num_code)-50)
    
  
data_period_plot <- ggplot(datasets_duration, aes( y=TIMESTAMP2,x=reorder(num_code,n))) +
                        geom_tile(fill = treeclass_pal[7])+
                        coord_flip()+
  theme_light()+
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+
guides(fill='none')+labs(x="",y='Year')+
  theme(axis.title.x=element_text(size=14),axis.text.x=element_text(size=12))

# Figure composition

cowplot::plot_grid(data_duration_plot, data_period_plot,labels=c('(a)','(b)'),rel_widths=c(1,0.4))


```

Figure 7. Datasets duration and period

\pagebreak

```{r fingerprint_plots, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, fig.height=18, fig.width=15}

# Data path
path.sapwood <- file.path('data/0.1.3/RData/sapwood')

# Plot
cowplot::plot_grid(
sfn_finger_species(read_sfn_data('ESP_CAN',folder=path.sapwood)),
sfn_finger_species(read_sfn_data('USA_WVF',folder=path.sapwood),years=1998,
                   species=c('Acer saccharum','Prunus serotina','Quercus alba','Quercus rubra')),

sfn_finger_species(read_sfn_data('CRI_TAM_TOW',folder=path.sapwood),years=2015,
                   species=c('Pouteria sp.','Inga sp.',
                                            'Eschweillera sp.','Mortoniodendron anisophyllum')),
ncol=1,labels=c('a)','b)','c)'),rel_heights = c(1,1,1))

```

Figure 8. Temporal patterns.

\pagebreak

Figure 9. Scaling



